\documentclass[11pt]{article}

\title{EMMA: Extensible MATLAB Medical Analysis \\ User Manual}

\author{Mark Wolforth \and Greg Ward \and Sean Marrett}
\date{20 November, 1995}

\def\code#1{{\tt \bf #1}}

\begin{document} 

\maketitle
\newpage

\tableofcontents

%--------------------------------------------------------------
\newpage
\section{Introduction}


This document is the user manual for EMMA (Extensible MATLAB Medical
Analysis) developed at the Montreal Neurological Institute.  This
package consists of a set of MATLAB .m files and programs written in
C, designed to ease the analysis of medical image data.  The entire
package runs under MATLAB 4.X.

This document assumes that the reader is familiar with manipulation of
medical images (specifically images stored in the MINC file
format)\footnote{MINC and the associated netCDF package are available
via anonymous FTP from ftp.bic.mni.mcgill.ca, in \code{/pub/minc}.},
and simply wishes to know how image manipulation is possible using EMMA
and MATLAB.

The authors would like to thank Sean Marrett of the McConnell Brain
Imaging Centre for being the driving force behind this project.  He
provided invaluable guidance when all seemed lost.

%--------------------------------------------------------------
\newpage
\section{Basic Concepts}

To a programmer, the interface to EMMA should be very familiar.
Simply open a data set contained in a MINC file by calling the
\code{openimage} function.  This returns a handle which is used in all
subsequent calls to EMMA functions.  The user does not need to worry
about the underlying mechanisms.  For example, to open an image file,
read in the the first frame of the second slice, and then close the
image file, you could type:

%
\begin{verbatim}
     data_handle = openimage ('arnaud_20547.mnc');
     image = getimages (data_handle, 2, 1); 
     closeimage (data_handle);
\end{verbatim}
%

In this example, \verb+data_handle+ is a handle to the data contained
in the MINC file \verb+arnaud_20547.mnc+.  After the data file has
been opened with \code{openimage}, the name of the MINC file is not
needed by the user of EMMA; only the handle for the data is used.
Therefore, \code{getimages} is passed the handle for the data, as well
as the slice number and frame number.  Once you have finished working
with the image, always remember to call \code{closeimage}.  This
cleans up global variables created in the MATLAB workspace.

Because EMMA uses handles, the user never needs to worry about the
interface between MATLAB and the MINC file.  You simply call EMMA
functions with the handle, and the underlying structure takes care of
the details.

However, the user {\em does} need to have some familiarity with
MATLAB, and a good understanding of how EMMA functions deal with
images.  In particular, the inherent two-dimensionality of MATLAB
means that a full four-dimensional MINC file is ``flattened'' in some
way when read into MATLAB.

Generally, EMMA functions deal with images transformed from their
natural two-dimensional form (well-suited to a matrix representation)
to a one-dimensional vector form.  For example, a single 128 $\times$
128 image will become a {\em column} vector of dimension 16384
$\times$ 1.  We can put several such vectors together in a matrix.
For example, we can form a 16384 $\times$ 15 matrix from 15 slices.
Thus, in an ``image matrix''---which is a two-dimensional MATLAB
matrix representing three-dimensional image data---the individual
columns correspond to whole images.  MATLAB functions that use EMMA to
perform analysis should always expect image data in this format, as
shown in Section
\ref{simple_example} and the document ``rCBF Analysis Using MATLAB.''

Also, keep in mind that EMMA makes it deliberately difficult to deal
simultaneously with the full four dimensions allowed in a MINC image
variable.  The reasoning for this is two-fold: first, collapsing three
dimensions into two via the ``image matrix'' concept is fairly
convenient and easy to work with; however, attempting to similarly
collapse four dimensions would add another layer of complications,
making both the implementation and use of EMMA functions more
difficult.  Second is the question of memory usage: MATLAB is capable
of using up enormous amounts of memory if care is not taken, and
restricting the user of EMMA to three image dimensions at a time means
that it is somewhat more difficult to blindly read in massive amounts
of image data with a single MATLAB command.

%--------------------------------------------------------------
\newpage
\section{Basic EMMA MATLAB Functions}

The basic MATLAB interface to EMMA is performed through the following
MATLAB functions (.m files).  This section is only meant to provide a
brief introduction to the most important functions.  See section
\ref{emma_reference} for full help on every EMMA function.

\begin{description}

\item \code{openimage} - Prepares a MINC file for reading.
\code{openimage} determines the image size, reads in the frame start
times and lengths if applicable, and returns a handle which can be
used to access the MINC file with \code{getimages},
\code{getimageinfo}, \code{getblooddata}, etc.
\begin{verbatim}
     handle = openimage (filename)
\end{verbatim}

\item \code{getimages} - Gets images from a MINC file.  If the file
has no frame dimension, the parameter \verb|frames| should be empty
or not provided.  For files with no slices (only frames), the
\verb|slices| parameter should be made empty (a matrix of the form
\verb|[]|).  Also, \code{getimages} can read several frames from a
single slice or several slices from a single frame---but it can not
read multiple slices and multiple frames simultaneously.  (This
restriction is also imposed for writing images, mainly to avoid
complications and reduce the software's ability to easily read in far
more data than can be handled by MATLAB.)

Two important features of \code{getimages} are the ability to re-use
memory, and the ability to read partial images.  Both of these
features are intended to reduce the amount of memory used by MATLAB
when processing images.
\begin{verbatim}
     Images = getimages (handle [, slices [, frames [, old_matrix ...
                         [, start_row [, num_rows]]]]])
\end{verbatim}

\item \code{newimage} - Creates a MINC file for a new set of images.
Returns a handle to the newly created data set that is used in calls
to other functions (such as \code{putimages} and \code{closeimage}).
\code{newimage} has a number of optional parameters, the most
important of which is the ``parent file.''  If this is supplied,
\code{newimage} will copy a number of important attributes (such as
the patient, study, and acquisition data) from the parent to the new
file, as well as using the parent file to provide the default image
size and type.
\begin{verbatim}
     handle = newimage (filename, dim_sizes [, parent_file [, image_type ...
                        [, valid_range [, orientation]]]]);
\end{verbatim}

\item \code{putimages} - Writes entire images to a file created with
\code{newimage}.  ({\bf Warning:} there are currently no provisions
made to deny the user from writing to a MINC file opened with
\code{openimage}.)  The syntax is similar to \code{getimages}, except
that the image data is of course an input argument to \code{putimages}
rather than an output argument.
\begin{verbatim}
     putimages (handle, images [, slices [, frames]])
\end{verbatim}

\item \code{getimageinfo} - Gets information about an open image.
Currently, \code{getimageinfo} will return the filename, number of
frames or slices, image height, width, or size, the dimension sizes,
frame start times, lengths, and mid-frame times.  The desired
information is specified as a character string, eg. 'NumFrames',
'ImageHeight', etc.
\begin{verbatim}
     info = getimageinfo (handle, info_descriptor)
\end{verbatim}

\item \code{viewimage} - Displays an image on the workstation screen.
On a monochrome display, the \code{gray} colourmap will be used;
otherwise, the \code{spectral} will be used.  The image will be scaled
such that the highest point in the image is always the last colour in
the colourmap, and the lowest image point will be the first element of
the colourmap.  Also, a colourbar relating colours in the image to the
values will be displayed, unless the optional \code{colourbar\_flag}
argument is set to zero.

\begin{verbatim}
     viewimage (image [, colourbar_flag])
\end{verbatim}

\item \code{closeimage} - Destroys the appropriate variables in the workspace.
\begin{verbatim}
     closeimage (IMhandle)
\end{verbatim}

\end{description}

\section{Some Basic Examples: Masking and Summing}
\label{simple_example}

As an example of the EMMA functions presented thus far, we will
explore sample code for both masking and summing/integrating images
from a dynamic study.  The MINC file used in this example will be one
of the H$_2$$^{15}$O CBF studies from
\code{/usr/local/matlab/toolbox/emma/examples}; there is also a
corresponding BNC ({\bf b}lood {\bf n}et{\bf C}DF) file, which is
necessary to get the plasma activity data.  To make accessing
these files easier, you may want to create symbolic links to them in a
directory of your own:
\begin{verbatim}
     cd <your-work-directory>
     ln -s /local/matlab/toolbox/emma/examples/yates_19445.* .
\end{verbatim}
That way, the file
\verb|/local/matlab/toolbox/emma/examples/yates_19445.mnc| can be
accessed as simply \verb|yates_19445.mnc| (as long as you are in your
work directory).  In the remainder of this section, it will be assumed
that these symbolic links exist in the current directory.

\subsection{Reading the Data}

First, let us open the MINC file:
\begin{verbatim}
     handle = openimage ('yates_19445.mnc');
\end{verbatim}
Now, we may wish to find out some basic data about the MINC file.  Try
the following (note that we will need \verb|nf|, the number of frames,
below):
\begin{verbatim}
     ns = getimageinfo (handle, 'NumSlices')
     nf = getimageinfo (handle, 'NumFrames')
     getimageinfo (handle, 'ImageSize')
\end{verbatim}
Note that for the study \verb|yates_19445|, \verb|nf| will be 21.
However, it is a good idea to always use \verb|getimageinfo| to find
the number of frames rather than assuming that it will always stay the
same for a particular type of study.

Now, let us read in the frame lengths and mid-frame times:
\begin{verbatim}
     FrameLengths = getimageinfo (handle, 'FrameLengths');
     MidFTimes = getimageinfo (handle, 'MidFrameTimes');
\end{verbatim}
Note that \verb|FrameLengths| and \verb|MidFTimes| are both column
vectors.  This will be important when performing matrix arithmetic
with them.

Now, we read in all frames for one slice---slice 8 in this case,
although you can work with whatever slice you like:
\begin{verbatim}
     slice = 8;
     PET = getimages (handle, slice, 1:nf);
\end{verbatim}
Note that {\em this} is the step where MATLAB uses up large amounts of
memory.  If you execute the MATLAB \verb|whos| command now, you will
see that MATLAB's variables alone are taking up nearly 3 megabytes of
memory.  However, the entire MATLAB workspace generally takes up
considerably more memory than is reported by \verb|whos|, because of
MATLAB's overhead and inefficient memory management scheme.

\subsection{Integrating Images}

We can view the matrix \verb|PET| as a collection of 21 PET images
for the same brain slice but progressing in time, or as a collection
of 16,384 time-activity curves (TAC)---one for every pixel in the 128
$\times$ 128 image.  (Of course, only about $\frac{1}{3}$ of the
image data is actually data from inside the head.  This matter will
be addressed when we consider image masking below.)

For purposes of integrating images, we will consider \verb|PET| as a
collection of TACs.  Thus, we are simply performing 16384 numerical
integrations.  This may seem a daunting task, but MATLAB's vectorized
approach to mathematics makes it surprisingly easy.  In fact, let us
calculate a rectangular integration using the formula:
\begin{equation}
\int_{t_{1}}^{t_{n}} f(t) dt \approx \sum_{i=1}^{n} f(t_{i}) (\Delta t)_{i},
\label{eq:rect_int}
\end{equation}
where $f(t)$ is one time-activity curve and the $\Delta t$'s are the
frame lengths.  Since we have $f(t)$ stored as a row
vector---remember, each column of \verb|PET| is the entire image for
one frame, so each row of \verb|PET| corresponds to a single pixel
across time---and $\Delta t$ stored as a column vector, the
multiply/sum operation of equation (\ref{eq:rect_int}) is simply a
dot product.  And since a matrix multiplication is nothing more than
a sequence of dot products, all 16,384 TACs can be quite easily
integrated with a single MATLAB command:
\begin{verbatim}
     PETsummed1 = PET * FrameLengths;
\end{verbatim}

However, rectangular integration is a very crude approach to
numerical integration.  Trapezoidal integration provides a slightly
more refined method, and is almost as easy to implement.
Furthermore, it is provided by the MATLAB function
\verb|trapz|---which, if you care to examine it (try 
\verb|type trapz| in MATLAB), is essentially a one-line operation as well.
(Incidentally, one of the functions provided by EMMA is
\code{ntrapz}.  \code{ntrapz} currently has exactly the same
restrictions and capabilities as \verb|trapz|, but is considerably
faster due to being written in C using MATLAB's C interface.  For the
purposes of this discussion, the two functions are equivalent.)

\verb|trapz| takes two arguments: a vector of $x$-points, and a vector
or matrix of $y$ points corresponding to each $x$-point.  If the $y$'s
are given as a matrix, then each {\em row} corresponds to one $x$
point.  Note that this is different from the EMMA standard, where each
{\em column} of \verb|PET| corresponds to one time point.  Thus, we
will have to transpose \verb|PET| using the MATLAB \verb|'| operator
before passing it to \verb|trapz|.  Also, \verb|trapz| will return the
16,384 integrals as a row vector, so we must transpose the results to
get them back into the form we expect.  Finally, \verb|trapz| expects
points on the $x$-axis rather than widths of intervals.  Thus, we will
pass \verb|MidFTimes| to it, and not \verb|FrameLengths|.  The full
integration is performed by the single statement
\begin{verbatim}
     PETsummed2 = trapz (MidFTimes, PET')';
\end{verbatim}

{\bf Note:} This method of calculating the integral is
extremely wasteful of memory due to transposing the PET images.  For a
full discussion of MATLAB memory issues, please consult the online
document Controlling MATLAB Memory
Use\footnote{http://www.mni.mcgill.ca/system/mni/matlab/matlab\_memory.html}.



\subsection{Viewing and Comparing the Integrated Images}

Now, we wish to view the integrated images.  If you are using a colour
console or an monochrome X terminal, enter the following to create a
MATLAB figure window and display the image obtained by rectangular
integration:
\begin{verbatim}
     figure;
     viewimage (PETsummed1);
\end{verbatim}
(Note that explicitly creating the window via the \verb|figure|
command is not really necessary.  However, if you already had a
figure window with a plot or image that you did not want to lose,
MATLAB would have overwritten the existing figure when you called
\code{viewimage}.  It is generally good practice to call
\verb|figure| before creating new plots or images, unless you are
sure you wish to obliterate whatever was previously in the current
figure window.)

You may want to title this window:
\begin{verbatim}
     title ('yates_19445, slice 8, rectangular integration');
\end{verbatim}

Now do the same thing for \verb|PETsummed2|.  You will probably want
to move one or both figure windows so that you can see both images
simultaneously.  Clearly, the two images are similar---but if we want
a more objective comparison, we can calculate the ratio of the two
images on a pixel-by-pixel basis and view this:
\begin{verbatim}
     ratio = PETsummed1 ./ PETsummed2;
     figure;
     viewimage (ratio);
\end{verbatim}

Note the use of the \verb|./| operator to specify an
element-by-element operation on two matrices.  The resulting image
shows that most of the pixels in \verb|PETsummed1| and
\verb|PETsummed2| appear fairly close.  However, because some pixels
have such large values compared to 1, any small variations around 1
are swamped.  One way to deal with this is simply to clip the image
at certain points: for example, set all points below -10 to -10, and
all points above 10 to 10.  This is accomplished as follows:
\begin{verbatim}
     clipneg = find (ratio < -10);
     ratio (clipneg) = ones (size (clipneg)) * (-10);
     clippos = find (ratio > 10);
     ratio (clippos) = ones (size (clippos)) * (10);
\end{verbatim}

If you wish to know how many points were clipped at either end,
type \verb|size (clipneg)| or \verb|size (clippos)|.

If you now \code{viewimage (ratio)}, it should be clear that the
overall ratio between the two integrated images is close to 1 inside
the head.

\subsection{Saving the Integrated Image}

After performing some analysis, it is often desirable to write the
results to a new MINC file.  Currently, we only have a single image to
write, so we will create a MINC file with 1 slice and no frames
(because of the integration across frames, time is no longer a
variable).  We use EMMA function \code{newimage} as follows:
\begin{verbatim}
     new_handle = newimage ('yates_summed.mnc', [0 1], ...
                            'examples/yates_19445.mnc');
\end{verbatim}
To write our single image to slice 1 of the new MINC file, we enter:
\begin{verbatim}
     putimages (new_handle, PETsummed2, 1);
\end{verbatim}

Alternately, we could create a MINC file with a full 15 slices, and
loop through the entire original study to sum every slice.  MATLAB
will allow us to enter a \verb|for| loop interactively, so enter the
following (note that if you already executed the above code, you
should use a new filename in place of \code{yates\_summed.mnc} or
delete the old one---\code{newimage} will die if you try to overwrite
an existing file):
\begin{verbatim}
     closeimage (new_handle);
     new_handle = newimage ('yates_summed.mnc', [0 ns], ...
                            'examples/yates_19445.mnc');
     for slice = 1:ns
        PET = getimages (handle, slice, 1:nf);
        PETsummed = trapz (MidFTimes, PET')';
        putimages (new_handle, PETsummed, slice);
     end
     closeimage (new_handle);
\end{verbatim}

\subsection{Using an Integrated Image for Image Masking}

First, let us clear some of the unneeded variables from the workspace:
\begin{verbatim}
     clear PETsummed1 ratio clipneg clippos
\end{verbatim}

Now, we will use \verb|PETsummed2| for a very simple form of image
masking: namely, we will use only pixels whose values in the
integrated image are greater than the mean of the integrated image.
For CBF studies, this usually creates a usable mask that contains only
in-brain voxels.  It can be done by calculating the mean of the
integrated image, and creating another image consisting entirely of
1's and 0's (a ``binary'' image), where a 1 indicates that the pixel
is inside the head, and a zero indicates outside:
\begin{verbatim}
     avg = mean (PETsummed2);
     mask = PETsummed2 > avg;
\end{verbatim}

We can view the mask as an image itself to see a silhouette of the
brain at this slice:
\begin{verbatim}
     figure;
     viewimage (mask);
\end{verbatim}
Or, we can apply the mask to the summed image and view it.  To apply
the mask to an image, simply perform an element-by-element multiply 
using the
\verb|.*| operator on the mask and the image; since \verb|mask|
consists entirely of ones and zeros, every point in the image will
either be set to zero or unchanged.  Thus,
\begin{verbatim}
     PETsummed2 = PETsummed2 .* mask;
     figure;
     viewimage (PETsummed2);
\end{verbatim}
will display the trapezoidally integrated image with out-of-head data
set to zero.

Finally, to see just how many pixels are ``on'' in the mask, you can type
\begin{verbatim}
     size (find (mask))
\end{verbatim}
The fact that only a fraction of the entire image consists of image
data means that masking can be used to great advantage when
performing analysis that requires computations for every pixel.
(Generally, this should be avoided if it is all possible to take
advantage of MATLAB's vectorized structure.)

%--------------------------------------------------------------
\newpage
\section{Underlying Structure}

EMMA has a set of underlying functions written in C that provide the
interface to MINC files.  An EMMA user will not ordinarily require
use of these functions since every interaction with MINC files should
be possible using the MATLAB scripts and functions provided (.m
files).  However, for the sake of completeness, these functions are
documented in this section.

MATLAB provides an interface to external C programs that allows
dynamic linking at runtime.  This interface, called CMEX, is detailed
in the MATLAB {\em External Interface Guide}.  Please consult this
guide for further details.  Unfortunately, due to problems with the
CMEX interface, it was not possible to create any CMEX programs that
wrote to MINC files.  Therefore, all of the MINC creation and writing
functions are written as stand-alone C programs, and all of the MINC
reading functions are written as CMEX programs.


\subsection {Stand-alone C Programs}

The following programs are used by various EMMA functions and are not
normally needed by the user.

\begin {itemize}

\item \code{miwriteimages} - Writes images to a MINC file.  The
images are taken from the specified temporary file, whose data is
expected to be stored as doubles.  The MINC file must exist, and
contain an image structure (create with micreate and then
micreateimage).
\begin{verbatim}
    miwriteimages <MINC file> <slices> <frames> <temp file>
\end{verbatim}

\item \code{micreateimage} - Create a new MINC file, complete with
image dimensions, dimension and dimension-width variables, image
max/min variables, and (of course) the image variable.  If the new
file is based on a ``parent'' MINC file, any other variables and
global attributes will be copied from the parent to the new file.  As
much information as {\em can} be copied about the image and image
dimensions will also be copied, but the fact that the new file may or
may not have the same orientation, dimensions, and dimension lengths
as its parent complicates matters greatly.
\begin{verbatim}
    micreateimage <MINC file> [option] [option] ...
\end{verbatim}

\item \code{miwritevar} - Writes values to a variable in a MINC file.  
Values are taken from the specified temporary file, whose data is
expected to be stored as doubles.
\begin{verbatim}
    miwritevar <MINC file> <var name> <start vector> ...
               <lengthvector> <temp file>
\end{verbatim}

\end{itemize}


\subsection{CMEX Programs}

In addition to stand-alone C programs, the underlying structure also includes
several CMEX programs which are dynamically linked by MATLAB.  Please see the
MATLAB {\em External Interface Guide} for an explanation of CMEX files.

\begin{itemize}

\item \code{mireadvar} - Read a hyperslab from a MINC (or any NetCDF)
file into a MATLAB vector, using NetCDF-style start and count
vectors.  Note that the indexing of \verb|start_vector| and
\verb|stop_vector| is zero-based.
\begin{verbatim}
    mireadvar ('minc_file', 'var_name', start_vector, count_vector)
\end{verbatim}

\item \code{mireadimages} - Read images from a MINC file.  Opens the
given MINC file, and attempts to read whole or partial images from the
slices and frames specified in the slices and frames vectors.  If
\verb|start_row| and \verb|num_rows| are not specified, then whole
images are read.  If only \verb|start_row| is specified, a single row
will be read.  If both \verb|start_rows| and \verb|num_rows| are
given, then the specified number of rows will be read (unless either
is out of bounds, which will result in an error message).

\begin{verbatim}
    mireadimages ('minc_file'[, slices[, frames[, options]]]) 
\end{verbatim}

\item \code{miinquire} - find out various things about a MINC file
from MATLAB.  Retrieves some item(s) of information about a MINC file.
The first argument is always the name of the MINC file.  Following the
filename can come any number of ``option sequences'', which consist of
the option (a string) followed by zero or more items (more strings).
Generally speaking, the ``option'' tells miinquire the general class
of information you're looking for (such as an attribute value or a
dimension length), and the item or items that follow it give
\code{miinquire} more details, such as the name of a dimension,
variable, or attribute.

\begin{verbatim}
    miinquire ('minc_file' [, 'option' [, 'item']], ...)
\end{verbatim}

\end{itemize}

%--------------------------------------
\newpage
\subsection{MATLAB files}
\label{emma_reference}

\input{emma-help}

\end{document}

